Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations

### 2018-01-18 ms.tex
more comments

\title{Global Climate and Atmosphere Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations}

@@

\begin{abstract}
	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC. The dayside planet-to-star flux is $0.151\pm0.015\%$, $0.446\pm0.38\%$, and $0.569\pm0.014\%$ in bandpasses centered at $1.4$, $3.6$, and $4.5\mu$m, respectively.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside. We fit the phase variation with a range of climate maps and find they all provide reasonable fits to the data, but have large degeneracies in temperature (up to 1000 K for a given latitude/longitude). The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with effective temperatures ranging from 2000 K on the nightside to 3000 K on the dayside. However, the dayside spectrum has a strong feature in the \Spitzer\ bands, which we attribute to CO emission and a thermal inversion.  The emission feature is not present on the nightside, suggesting changing thermal structure with longitude. We retrieved the atmospheric composition with a 1D thermochemical equilibrium model and found the composition is moderately metal-enriched ($\mathrm{[M/H]} =  23^{+29}_{-13}\times$ solar), and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to other hot Jupiters, we do not detect significant water absorption, which we attribute to partial H$_2$O dissociation and additional opacity from H- and TiO/VO.  We compare the phase curves to 3D general circulation models and find that atmospheric drag is required to explain the small hotspot offsets, but that these models overpredict the phase curve amplitude.  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find they have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.
\end{abstract}

@@

\subsection{\HST/WFC3}
The \HST\ phase curve observations consisted of two visits on 26-27 February and 2-3 August 2015. Each visit was 15 orbits in duration and spanned 23 hours. The last half of orbit 15 in each visit was used for a gyro bias update and does not produce useable science data.  We took a direct image of the star with the F126N filter at the beginning of each orbit to determine the wavelength solution zero-point. The remainder of the orbit consisted of time-series spectroscopy with the G141 grism ($1.1 - 1.7$ $\mu$m) and the 256 x 256 pixel subarray. We used the SPARS10/NSAMP = 15 read-out mode, which has an exposure time of 103 seconds. To optimize the duty cycle of the observations, we used the spatial scan observing mode with a scan rate of 0.03 arcsec/s, alternating between forward and backward scanning on the detector. The scan height was 25 pixels and the peak counts were $35\times10^3$ photoelectrons per pixel. We collected a total of 18 spatial scan exposures per orbit.  The two eclipse observations from Program 13660 had a similar observing setup \citep[described in detail in][]{cartier17}.

@@

The data were reduced with the POET pipeline \citep{stevenson12}. We performed aperture photometry with an aperture size of 2.75 pixels (chosen from a grid of apertures between 2 and 4 pixels to minimize the residual noise in the light curve fits). We masked pixels that were flagged in the bad pixel mask provided in the ancillary data for the observations. The target centroid was determined with a two-dimensional Gaussian fit.  We estimated and subtracted the background from an annulus with a radius of 7 to 15 pixels from the centroid. As for the \HST\ observations, the contaminating flux from the nearby star is included in the final photometry and corrected later in the light curve fits.

@@

The photometric analyses are summarized for each observing season in Table\,\ref{tab:photometry}.  The standard deviation of WASP-103's brightness in each season is given in column~4.  The mean of the three standard deviations is 0.0058~mag, which is very comparable to the mean standard deviation of the six comparison stars (0.0055~mag).  To maximize the possibility of detecting WASP-103b's rotation, we normalized the photometry such that each observing season has the same mean, thereby removing any long-term variability in WASP-103 and/or the comparison stars (Figure\,\ref{fig:photometry}, top panel).  We performed a periodogram analysis of the normalized data set based on least-squares fitting of sine curves.  The resulting frequency spectrum and the phase curve computed with the best period are shown in the middle and lower panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is 6.814 days, which agrees closely with the estimated stellar rotation period of 6.855 days of \citep{gillon14}. The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,ppm from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

We calculated the planet-to-star flux $F_p/F_s$ in two different ways. First, we fit a sinusoid with a period equal to the planet's orbital period. The free parameters were the sine curve amplitude and phase offset. For the second approach, we used the \texttt{spiderman} package \citep{louden17} to model $F_p/F_s$. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding flux ratio for an observation at time $t$. In our fit, we calculated the stellar flux with a PHOENIX model \citep{husser13} interpolated to an effective temperature of $6110\,\mathrm{K}$ \citep{gillon14}.  For the planet flux, we tested three different maps: a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang17}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$).  In all cases, we assumed that the planet is tidally locked, such that each orbital revolution corresponds to one complete rotation on its spin axis.

@@

\subsection{Goodness of Fit}
\label{sec:fitquality}
We performed several tests of the quality of the light curve fits.  First we predicted the level of scatter in the light curves based on photon noise alone, then compared this value to the root-mean-square (rms) of the fit residuals.  The Spitzer 4.5 $\mu$m light curve rms is within 10\% of the expected photon noise limit (691 versus 640 ppm), whereas the 3.6 $\mu$m light curve has significantly larger rms (797 versus 470 ppm), due to time-correlated red noise (discussed below). The expected photon-limited rms for the WFC3 spectrosopic light curves ranges from 430 - 530 parts per million (ppm), and the measured rms was typically within 5\% of expectations for all spectroscopic channels.  For the WFC3 band-integrated white light curve, the rms was slightly larger than predicted (175 versus 122 ppm). There are a number of possible origins for this discrepancy, including imperfect background subtraction, variation in the position of the spatial scan on the detector, loss of flux outside the extraction window, and imperfect modeling of the astrophysical signal or instrument systematics. However, our analysis of the composition and thermal structure of the planet's atmosphere is based on the wavelength-dependent data (where we reach the photon limit), so we do not further explore the small excess noise in the white light curve.

@@

For all three phase curves, the hotspot offset is consistent with zero degrees, indicative of a fast radiative-to-advective timescale (the incident flux is reradiated to space faster than it is advected around to the nightside). Fast radiative timescales are precicted at high temperatures, and small hotspot offsets are also observed for other very hot Jupiters \citep{komacek17}.  The phase curve amplitudes are large (near 0.8 - 0.9), as expected for an atmosphere with inefficient heat redistribution. In \S\,\ref{sec:gcm} we compare these results to expectations from 3D GCMs.

@@

\begin{deluxetable}{llLL}
	\tablecolumns{4}
	\tablewidth{0pt}
	\tablecaption{Phase Curve Properties \label{table:amps_offsets}}
	\tablehead{
		\colhead{Bandpass} & \colhead{Source} & \colhead{Amplitude} & \colhead{Offset}\\
	\colhead{\,} & \colhead{\,} & \colhead{\,} & \colhead{(Degrees)}}
		\startdata
		WFC3 & data & 0.91 \pm 0.02 & -0.3 \pm 0.1 \\
		\, & nominal GCM & 0.89 & 15.32 \\
		\, & [Fe/H] = 0.5 GCM & 0.84 & 19.64 \\
		\, & $\tau_\mathrm{drag4}$ GCM & 0.97 & 2.34 \\
		\, & $\tau_\mathrm{drag3}$ GCM & 0.99 & 0.18 \\
		Spitzer 3.6 $\mu$m & data & 0.86 \pm 0.13 & 2.0 \pm 0.7 \\
		\, & nominal GCM & 0.78 & 9.19 \\
		\, & [Fe/H] = 0.5 GCM & 0.72 & 12.79 \\
		\, & $\tau_\mathrm{drag4}$ GCM & 0.86 & 0.90 \\
		\, & $\tau_\mathrm{drag3}$ GCM & 0.97 & 0.18 \\
		Spitzer 4.5 $\mu$m & data & 0.83 \pm 0.05 & 1.0 \pm 0.4 \\
		\, & nominal GCM & 0.79 & 8.11 \\
		\, & [Fe/H] = 0.5 GCM & 0.73 & 11.35 \\
		\, & $\tau_\mathrm{drag4}$ GCM & 0.85 & 0.90 \\
		\, & $\tau_\mathrm{drag3}$ GCM & 0.93 & 0.18 \\
		\enddata
	\end{deluxetable}

@@

To test that the phase-resolved spectra are robust to different approaches for fitting the data, we compared the spectra for all four of the thermal phase variation models (sinusoid, kinematic, spherical harmonics, and two temperature). We found that the choice of model does not significantly change the results.  The emission spectra are insensitive to the models because they are estimated directly from the data (after systematics have been removed). Since systematic noise is not strongly correlated with the astrophysical signal, the systematics-divided data are nearly identical for all the models.  This point is illustrated in Figure\,\ref{fig:model_comparison} for the broadband WFC3 light curve. In most phase bins, the planet-to-star flux agrees to well within one sigma for all the models.

@@

We compared the planet's minimum and maximum temperature for the three climate maps by computing the temperatures over a $100\times100$ grid in latitude and longitude. The kinematic model has the hottest dayside temperature, and the two-temperature model has the smallest. The differences in peak temperature for the models we consider ranges from $630\,\mathrm{K}$ for \Spitzer\ Channel 1 to $1070\,\mathrm{K}$ for the broadband WFC3 light curve. The minimum nightside temperature is also model-dependent: for the WFC3 data, the two-temperature model predicts a nightside temperature of zero Kelvin, whereas the best fit kinematic model has a nightside temperature of $1980\,\mathrm{K}$.  These differences arise because the kinematic model allows a steep temperature gradient on the dayside, so the substellar point is much hotter than the terminator, whereas the two temperature map imposes a constant dayside temperature.  The spherical harmonics model has an intermediate temperature gradient and fits the data the best. However, this model may not be physically realistic: on the nightside, it produces higher temperatures at the poles than at the equator, contrary to predictions from GCMs (which we discuss in \S\,\ref{sec:gcm}).

@@

\subsection{Comparison with Previous Observations}
We compared our results to the dayside emission spectrum reported by \cite{cartier17}, which is based on a Gaussian process analysis of two secondary eclipses from \HST/WFC3. The shape of their spectrum is consistent with what we find, but their eclipse depths are 125 ppm smaller on average. In their analysis, \cite{cartier17} noted that the Gaussian process fit produces smaller eclipse depths than a parametric fit (also by 125 ppm). One explanation for this difference is that the Gaussian process models some of the astrophysical signal. If the planet's thermal phase variation is absorbed into the instrument systematic model, the measured eclipse depths would be lower by about 100 ppm. Since our analysis accounts for the thermal phase variation, it is more likely to provide a robust estimate of the secondary eclipse depths.  Our estimated uncertainties are a factor of four smaller than those reported in \cite{cartier17}, which is consistent with photon-limited expectations (our data set includes two additional eclipses, a factor of five longer out-of-eclipse baseline, and has 60\% larger wavelength bins).

@@

We also fit blackbody models to the WFC3 data only, and found that they provide a good fit at all orbital phases. The $\chi^2_\nu$ values range from 0.5 to 1.6, with a median of 0.9. The best fit effective temperatures range from 1920 Kelvin on the nightside to 2980 Kelvin on the dayside. We report the best fit temperatures with uncertainties in Table\,\ref{table:teffs}.  The dayside spectrum is the poorest fit by the blackbody model. This finding is not a surprise because the dayside has higher signal-to-noise than the other orbital phases, thanks to the two secondary eclipse observations from \cite{cartier17}. The more sophisticated grid-based retrieval (described in the next section) provides a better fit to the dayside.

\begin{deluxetable}{LLL}
	\tablecolumns{3}
	\tablewidth{0pt}
	\tablecaption{Phase-resolved Effective Temperature \label{table:teffs}}
	\tablehead{
	\colhead{Orbital Phase} & \colhead{$T_\mathrm{eff}$ (K)} & \colhead{$\chi^2_\nu$}}
		\startdata
		0.06 - 0.15 & 1919 \pm 40 & 0.6 \\
		0.15 - 0.25 & 2256 \pm 24 & 0.8 \\
		0.25 - 0.35 & 2651 \pm 17 & 1.2 \\
		0.35 - 0.44 & 2906 \pm 11 & 0.8 \\
		0.44 - 0.56 & 3013 \pm 8 & 1.6 \\
		0.56 - 0.65 & 2885 \pm 12 & 1.3 \\
		0.65 - 0.75 & 2636 \pm 16 & 0.9 \\
		0.75 - 0.85 & 2313 \pm 23 & 1.1 \\
		0.85 - 0.94 & 2027 \pm 36 & 0.5 \\
		\enddata
		\vspace{-0.8cm}
		\tablecomments{comments}
	\end{deluxetable}

@@

The retrieved C/O is consistent with solar, with a $1\sigma$ confidence interval of $0.54 - 0.85$. We infer an upper limit on C/O of 0.9 at $3\,\sigma$ confidence, driven by the fact that the atmospheric chemistry is expected to change dramatically when C/O exceeds unity. For a carbon-rich composition, the equilibrium abundance of methane relative to CO increases by orders of magnitude compared to an oxygen-rich  composition \citep[e.g.][]{madhusudhan11}. Our \Spitzer\ eclipse depths are sensitive to the relative abundance of these species, so we can confidently rule out a carbon-rich composition despite the lack of spectrally resolved features.

@@

\section{Comparison with GCMs}
\label{sec:gcm}
To explore the three-dimensional effects of atmospheric dynamics, we ran several GCM models to compare with the measured phase-resolved spectra. (\textbf{Vivien: add text describing GCM}).  Our nominal GCM was a cloud-free, solar composition atmosphere with TiO/VO opacity and no added drag.  We also ran models with enhanced metallicity ([Fe/H] = 0.5), no TiO/VO, and added atmospheric drag.  Lorentz drag can occur in the hottest exoplanet atmospheres, which may be partially ionized and coupled to the background magnetic field. We parameterized the drag (\textbf{Vivien: how did you model drag?}), with timescales of $\tau_\mathrm{drag} = 10^3$ and $10^4$ s. We label the drag models $\tau_\mathrm{drag3}$ and $\tau_\mathrm{drag4}$, respectively.

@@

We selected spectra from WASP-103b at three orbital phases: dayside ($\phi = 0.5$), quadrature ($\phi = 0.25$), and nightside ($\phi = 0.1$), and compared them to sources with comparable effective temperature. Figure\,\ref{fig:planetstarcomparison} shows the flux-calibrated spectra (assuming a distance of 10 pc).  The directly imaged spectra are for the sources CD-35 2722, USco 1610-1913B, and TWA 22A and are taken from \cite{wahhaj11, aller13, bonnefoy14}. They are calibrated in absolute flux using published H-band photometry \citep{wahhaj11, aller13, bonnefoy09} and distances \citep{gaia16, teixeira09}, a flux-calibrated spectrum of Vega \citep{1985A&A...151..399M, 1985IAUS..111..225H}, and the corresponding filter passbands.  The brown dwarf spectra \textbf{Jackie: describe catalog of brown dwarf spectra and flux calibration}.

@@

We compared the spectra over the wavelength range $1.1 - 1.7\mu$m. The dominant spectral features over this range are expected to come from water, which has a forest of absorption lines near 1.4\,$\mu$m. Spectra for the brown dwarfs and imaged companions have noticeable features in the water band, whereas WASP-103b does not. To quantitatively compare the water feature amplitude for different objects, we define an amplitude $A = (F_{1,2} - F_{3,4})/F_{1,2}$, where $F_{1,2}$ is the weighted mean flux in a wavelength bin $\lambda_1 - \lambda_2$. We calculated the water feature amplitude for two choices of wavelength bins. For the first, $A_1$, we considered data in and out of the water band, with $\lambda_{1,2,3,4} = 1.15, 1.3, 1.35, 1.5\,\mu$m. The ground-based direct imaging data doesn't span this entire wavelength range, so we also define an amplitude $A_2$ with $\lambda_{1,2,3,4} = 1.2, 1.35, 1.5, 1.65\,\mu$m. The estimated amplitudes and uncertainties are listed in Table\,\ref{table:sources}.

@@

	\item{The dayside planet-to-star flux is $0.151\pm0.015\%$, $0.446\pm0.38\%$, and $0.569\pm0.014\%$ in the WFC3 bandpass, \Spitzer\ $3.6$, and \Spitzer\ $4.5\mu$m, respectively.  The best fit blackbody to the WFC3 dayside spectrum has an effective temperature of $3013 \pm 8$ K, making WASP-103b among the hottest hot Jupiters ever observed.}

>>>>

\title{Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations}

@@

\begin{abstract}
	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC. The dayside planet-to-star flux is $0.151\pm0.015\%$, $0.446\pm0.38\%$, and $0.569\pm0.014\%$ in bandpasses centered at $1.4$, $3.6$, and $4.5\mu$m, respectively.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside. We fit the phase variation with a range of climate maps and find they all provide reasonable fits to the data, but have large degeneracies in temperature (up to 1000 K for a given latitude/longitude). The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with brightness temperatures ranging from 2000 K on the nightside to 3000 K on the dayside. However, the dayside spectrum has a strong feature in the \Spitzer\ bands, which we attribute to CO emission and a thermal inversion.  The emission feature is not present on the nightside, suggesting changing thermal structure with longitude. We retrieved the atmospheric composition with a 1D thermochemical equilibrium model and found the composition is moderately metal-enriched ($\mathrm{[M/H]} =  23^{+29}_{-13}\times$ solar), and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to other hot Jupiters, we do not detect significant water absorption, which we attribute to partial H$_2$O dissociation and additional opacity from H- and TiO/VO.  We compare the phase curves to 3D general circulation models and find that atmospheric drag is required to explain the small hotspot offsets, but that these models overpredict the phase curve amplitude.  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find they have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.
\end{abstract}

@@

\subsection{\HST/WFC3}
The \HST\ phase curve observations consisted of two visits on 26-27 February and 2-3 August 2015. Each visit was 15 orbits in duration and spanned 23 hours. The last half of orbit 15 in each visit was used for a gyro bias update and does not produce useable science data.  We took a direct image of the star with the F126N filter at the beginning of each orbit to determine the wavelength solution zero-point. The remainder of the orbit consisted of time-series spectroscopy with the G141 grism ($1.1 - 1.7$ $\mu$m) and the 256 x 256 pixel subarray. We used the SPARS10/NSAMP = 15 read-out mode, which has an exposure time of 103 seconds. To optimize the duty cycle of the observations, we used the spatial scan observing mode with a scan rate of 0.03 arcsec/s, alternating between forward and backward scanning on the detector. The scan height was 25 pixels and the peak counts were $3.5\times10^4$ photoelectrons per pixel. We collected a total of 18 spatial scan exposures per orbit.  The two eclipse observations from Program 13660 had a similar observing setup \citep[described in detail in][]{cartier17}.

@@

The data were reduced with the POET pipeline \citep{stevenson12}. We performed aperture photometry with an aperture size of 2.75 pixels (chosen from a grid of apertures between 2 and 4 pixels to minimize the residual noise in the light curve fits). We masked pixels that were flagged in the bad pixel mask provided in the ancillary data for the observations. The target centroid was determined with a two-dimensional Gaussian fit.  The target remains centered near the sweet spot for the entire AOR in each observation, with a maximum drift of 0.1 pixels. We estimated and subtracted the background from an annulus with a radius of 7 to 15 pixels from the centroid. As for the \HST\ observations, the contaminating flux from the nearby star is included in the final photometry and corrected later in the light curve fits.

@@

The photometric analyses are summarized for each observing season in Table\,\ref{tab:photometry}.  The standard deviation of WASP-103's brightness in each season is given in column~4.  The mean of the three standard deviations is 0.0058~mag, which is very comparable to the mean standard deviation of the six comparison stars (0.0055~mag).  To maximize the possibility of detecting WASP-103b's rotation, we normalized the photometry such that each observing season has the same mean, thereby removing any long-term variability in WASP-103 and/or the comparison stars (Figure\,\ref{fig:photometry}, top panel).  We performed a periodogram analysis of the normalized data set based on least-squares fitting of sine curves.  The resulting frequency spectrum and the phase curve computed with the best period are shown in the middle and lower panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is 6.814 days, which agrees closely with the estimated stellar rotation period of 6.855 days of \citep{gillon14}. The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,parts per million (ppm) from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

We calculated the planet-to-star flux $F_p/F_s$ in two different ways. First, we fit a sinusoid with a period equal to the planet's orbital period. The free parameters were the sine curve amplitude and phase offset. For the second approach, we used the \texttt{spiderman} package \citep{louden17} to model $F_p/F_s$. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding flux ratio for an observation at time $t$. In our fit, we calculated the stellar flux with a PHOENIX model \citep{husser13} interpolated to an brightness temperature of $6110\,\mathrm{K}$ \citep{gillon14}, solar metallicity, and log\,$g$ of 4.5 (in cgs units).  For the planet flux, we tested three different maps: a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang17}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$).  In all cases, we assumed that the planet is tidally locked, such that each orbital revolution corresponds to one complete rotation on its spin axis.

@@

\subsection{Goodness of Fit}
\label{sec:fitquality}
We performed several tests of the quality of the light curve fits.  First we predicted the level of scatter in the light curves based on photon noise alone, then compared this value to the root-mean-square (rms) of the fit residuals.  The Spitzer 4.5 $\mu$m light curve rms is within 10\% of the expected photon noise limit (691 versus 640 ppm), whereas the 3.6 $\mu$m light curve has significantly larger rms (797 versus 470 ppm), due to time-correlated red noise (discussed below). The expected photon-limited rms for the WFC3 spectrosopic light curves ranges from 430 - 530 ppm, and the measured rms was typically within 5\% of expectations for all spectroscopic channels.  For the WFC3 band-integrated white light curve, the rms was slightly larger than predicted (175 versus 122 ppm). There are a number of possible origins for this discrepancy, including imperfect background subtraction, variation in the position of the spatial scan on the detector, loss of flux outside the extraction window, and imperfect modeling of the astrophysical signal or instrument systematics. However, our analysis of the composition and thermal structure of the planet's atmosphere is based on the wavelength-dependent data (where we reach the photon limit), so we do not further explore the small excess noise in the white light curve.

@@

For all three phase curves, the hotspot offset is consistent with zero degrees, indicative of a small ratio of radiative to advective timescales (the incident flux is reradiated to space faster than it is advected around to the nightside). Fast radiative timescales are precicted at high temperatures, and small hotspot offsets are also observed for other very hot Jupiters \citep{perez13, komacek16, komacek17}.  The phase curve amplitudes are large (near 0.8 - 0.9), as expected for an atmosphere with inefficient heat redistribution. In \S\,\ref{sec:gcm} we compare these results to expectations from 3D GCMs.

@@

\begin{deluxetable}{llLL}
	\tablecolumns{4}
	\tablewidth{0pt}
	\tablecaption{Phase Curve Properties \label{table:amps_offsets}}
	\tablehead{
		\colhead{Bandpass} & \colhead{Source} & \colhead{Amplitude} & \colhead{Offset}\\
	\colhead{\,} & \colhead{\,} & \colhead{\,} & \colhead{(Degrees)}}
		\startdata
		WFC3 & data & 0.91 \pm 0.02 & -0.3 \pm 0.1 \\
		\, & nominal GCM & 0.89 & 15.32 \\
		\, & [M/H] = 0.5 GCM & 0.84 & 19.64 \\
		\, & $\tau_\mathrm{drag4}$ GCM & 0.97 & 2.34 \\
		\, & $\tau_\mathrm{drag3}$ GCM & 0.99 & 0.18 \\
		Spitzer 3.6 $\mu$m & data & 0.86 \pm 0.13 & 2.0 \pm 0.7 \\
		\, & nominal GCM & 0.78 & 9.19 \\
		\, & [M/H] = 0.5 GCM & 0.72 & 12.79 \\
		\, & $\tau_\mathrm{drag4}$ GCM & 0.86 & 0.90 \\
		\, & $\tau_\mathrm{drag3}$ GCM & 0.97 & 0.18 \\
		Spitzer 4.5 $\mu$m & data & 0.83 \pm 0.05 & 1.0 \pm 0.4 \\
		\, & nominal GCM & 0.79 & 8.11 \\
		\, & [M/H] = 0.5 GCM & 0.73 & 11.35 \\
		\, & $\tau_\mathrm{drag4}$ GCM & 0.85 & 0.90 \\
		\, & $\tau_\mathrm{drag3}$ GCM & 0.93 & 0.18 \\
		\enddata
	\end{deluxetable}

@@

To test that the phase-resolved spectra are robust to different approaches for fitting the data, we compared the spectra for all four of the thermal phase variation models (sinusoid, kinematic, spherical harmonics, and two temperature). We found that the choice of model does not significantly change the results.  The emission spectra are insensitive to the models because they are estimated directly from the data (after systematics have been removed). Since the systematic noise is not strongly correlated with the astrophysical signal, the systematics-divided data are nearly identical for all the models.  This point is illustrated in Figure\,\ref{fig:model_comparison} for the broadband WFC3 light curve. In most phase bins, the planet-to-star flux agrees to well within one sigma for all the models.

@@

We compared the planet's minimum and maximum temperature for the three climate maps by computing the temperatures over a $100\times100$ grid in latitude and longitude. The kinematic model has the hottest dayside temperature, and the two-temperature model has the smallest. The differences in peak temperature for the models we consider ranges from $630\,\mathrm{K}$ for \Spitzer\ Channel 1 to $1070\,\mathrm{K}$ for the broadband WFC3 light curve. The minimum nightside temperature is also model-dependent: for the WFC3 data, the two-temperature model predicts a nightside temperature of zero Kelvin, whereas the best fit kinematic model has a nightside temperature of $1980\,\mathrm{K}$.  These differences arise because the kinematic model allows a steep temperature gradient on the dayside, so the substellar point is much hotter than the terminator, whereas the two temperature map imposes a constant dayside temperature.  The spherical harmonics model has an intermediate temperature gradient and fits the data the best. However, this model may not be physically realistic: on the nightside, it produces much higher temperatures at the poles than at the equator, contrary to predictions from GCMs (which we discuss in \S\,\ref{sec:gcm}).

@@

\subsection{Comparison with Previous Observations}
We compared our results to the dayside emission spectrum reported by \cite{cartier17}, which is based on a Gaussian process analysis of two secondary eclipses from \HST/WFC3. The shape of their spectrum is consistent with what we find, but their eclipse depths are 125 ppm smaller on average (a difference of about 10\%). In their analysis, \cite{cartier17} noted that the Gaussian process fit produces smaller eclipse depths than a parametric fit (also by 125 ppm). One explanation for this difference is that the Gaussian process models some of the astrophysical signal. If the planet's thermal phase variation is absorbed into the instrument systematic model, the measured eclipse depths would be lower by about 100 ppm. Since our analysis accounts for the thermal phase variation, it is more likely to provide a robust estimate of the secondary eclipse depths.  Our estimated uncertainties are a factor of four smaller than those reported in \cite{cartier17}, which is consistent with photon-limited expectations (our data set includes two additional eclipses, a factor of five longer out-of-eclipse baseline, and has 60\% larger wavelength bins).

@@

We also fit blackbody models to the WFC3 data only, and found that they provide a good fit at all orbital phases. The $\chi^2_\nu$ values range from 0.5 to 1.6, with a median of 0.9. The best fit brightness temperatures range from 1920 Kelvin on the nightside to 2980 Kelvin on the dayside. We report the best fit temperatures with uncertainties in Table\,\ref{table:teffs}.  The dayside spectrum is the poorest fit by the blackbody model. This finding is not a surprise because the dayside has higher signal-to-noise than the other orbital phases, thanks to the two secondary eclipse observations from \cite{cartier17}. The more sophisticated grid-based retrieval (described in the next section) provides a better fit to the dayside.

\begin{deluxetable}{LLL}
	\tablecolumns{3}
	\tablewidth{0pt}
	\tablecaption{Phase-resolved Brightness Temperature \label{table:teffs}}
	\tablehead{
	\colhead{Orbital Phase} & \colhead{$T_\mathrm{b}$ (K)} & \colhead{$\chi^2_\nu$}}
		\startdata
		0.06 - 0.15 & 1919 \pm 40 & 0.6 \\
		0.15 - 0.25 & 2256 \pm 24 & 0.8 \\
		0.25 - 0.35 & 2651 \pm 17 & 1.2 \\
		0.35 - 0.44 & 2906 \pm 11 & 0.8 \\
		0.44 - 0.56 & 3013 \pm 8 & 1.6 \\
		0.56 - 0.65 & 2885 \pm 12 & 1.3 \\
		0.65 - 0.75 & 2636 \pm 16 & 0.9 \\
		0.75 - 0.85 & 2313 \pm 23 & 1.1 \\
		0.85 - 0.94 & 2027 \pm 36 & 0.5 \\
		\enddata
		\vspace{-0.8cm}
		\tablecomments{comments}
	\end{deluxetable}

@@

The retrieved C/O is consistent with solar, with a $1\sigma$ confidence interval of $0.54 - 0.85$. We infer an upper limit on C/O of 0.9 at $3\,\sigma$ confidence, driven by the fact that the atmospheric chemistry is expected to change dramatically when C/O exceeds unity. For a carbon-rich composition, the equilibrium abundance of methane relative to CO increases by orders of magnitude compared to an oxygen-rich  composition \citep[e.g.][]{madhusudhan11}. Our \Spitzer\ eclipse depths are sensitive to the relative abundance of these species, so we can confidently rule out a carbon-rich composition despite the lack of spectrally resolved features (assuming the atmosphere is in chemical equilibrium).

@@

\section{Comparison with GCMs}
\label{sec:gcm}
To explore the three-dimensional effects of atmospheric dynamics, we ran several GCM models to compare with the measured phase-resolved spectra. (\textbf{Vivien: add text describing GCM}).  Our nominal GCM was a cloud-free, solar composition atmosphere with TiO/VO opacity and no added drag.  We also ran models with enhanced metallicity ([M/H] = 0.5), no TiO/VO, and added atmospheric drag.  Lorentz drag can occur in the hottest exoplanet atmospheres, which may be partially ionized and coupled to the background magnetic field. We parameterized the drag (\textbf{Vivien: how did you model drag?}), with timescales of $\tau_\mathrm{drag} = 10^3$ and $10^4$ s. We label the drag models $\tau_\mathrm{drag3}$ and $\tau_\mathrm{drag4}$, respectively.

@@

We selected spectra from WASP-103b at three orbital phases: dayside ($\phi = 0.5$), quadrature ($\phi = 0.25$), and nightside ($\phi = 0.1$), and compared them to sources with comparable brightness temperature. Figure\,\ref{fig:planetstarcomparison} shows the flux-calibrated spectra (assuming a distance of 10 pc).  The directly imaged spectra are for the sources CD-35 2722, USco 1610-1913B, and TWA 22A and are taken from \cite{wahhaj11, aller13, bonnefoy14}. They are calibrated in absolute flux using published H-band photometry \citep{wahhaj11, aller13, bonnefoy09} and distances \citep{gaia16, teixeira09}, a flux-calibrated spectrum of Vega \citep{1985A&A...151..399M, 1985IAUS..111..225H}, and the corresponding filter passbands.  The brown dwarf spectra \textbf{Jackie: describe catalog of brown dwarf spectra and flux calibration}.

@@

We compared the spectra over the wavelength range $1.1 - 1.7\mu$m. The dominant spectral features over this range are expected to come from water, which has a forest of absorption lines near 1.4\,$\mu$m. Spectra for the brown dwarfs and imaged companions have noticeable features in the water band, whereas WASP-103b does not. To quantitatively compare the water feature amplitude for different objects, we define an amplitude $A = (F_{1,2} - F_{3,4})/F_{1,2}$, where $F_{1,2}$ is the weighted mean flux in a wavelength bin $\lambda_1 - \lambda_2$. We calculated the water feature amplitude for two choices of wavelength bins. For the first, $A_1$, we considered data in and out of the water band, with $\lambda_{1,2,3,4} = 1.15, 1.3, 1.35, 1.5\,\mu$m. The ground-based direct imaging data do not span this entire wavelength range, so we also define an amplitude $A_2$ with $\lambda_{1,2,3,4} = 1.2, 1.35, 1.5, 1.65\,\mu$m. The estimated amplitudes and uncertainties are listed in Table\,\ref{table:sources}.

@@

	\item{The dayside planet-to-star flux is $0.151\pm0.015\%$, $0.446\pm0.38\%$, and $0.569\pm0.014\%$ in the WFC3 bandpass, \Spitzer\ $3.6$, and \Spitzer\ $4.5\mu$m, respectively.  The best fit blackbody to the WFC3 dayside spectrum has an brightness temperature of $3013 \pm 8$ K, making WASP-103b among the hottest hot Jupiters ever observed.}
